---
title: "Using Tab Function(s)"
author: "Jesse Zlotoff"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Tab Function(s)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#suppressPackageStartupMessages(library(rtabsplus))
library(devtools)
load_all()
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
suppressPackageStartupMessages(tab(iris,Species))
```

The rtabsplus package includes multiple versions of the `tab` function.

## One-way Tab, Unweighted

The most basic usage is to tabulate proportions of a factor variable.  Using the **iris** dataset as an example:

```{r}
# tab function
#   usage: tab(df, v1) where df is a dataframe and v1 is the column name
tab(iris, Species)
```

This provides similar functionality as the built-in `table` function with a few key differences.  It gives column proportions instead of case sizes and returns the results as a tibble.  You can use the optional `nsize` parameter to include those values as well.

```{r}
# tab function
tab(iris, Species, nsize=TRUE)

# compare to built-in table function
table(iris$Species)
```

There is also an option to include the upper and lower bounds of the 95% confidence interval.  Although this is not survey data, it is calculated here for demonstration purposes.

```{r}
# tab function
tab(iris, Species, nsize=TRUE, ci=TRUE)
```

The tab function also takes an optional parameter to create collapsed categories.  Each new category is specified in a list as "new label"=c(inputs).  Use "@auto" at the start of a label to auto-generate it.

```{r}
# tab function
my_collapses = list("@auto" = c("versicolor","virginica"), "@auto2" = c("setosa"))
tab(iris, Species, collapses=my_collapses)
```

## Two-way Tabs, Unweighted

The `tab` function can also handle two-way tabulations by specifying the optional `v2` parameter.

```{r}
# tab function
#   usage: tab(df, v1, v2=v2)
tab(iris, Species, v2=Petal.Width)
```

Again, the results are a tibble with column proportions, including Total.  Additional columns such as "n_" (for nsize) or "lci_", "uci_" (for ci) use the standard prefixes before the specific column name.

Note, the order of the variables matters.

```{r}
# Species vs Petal.Width
#   usage: tab(df, v1, v2=v2, nsize=FALSE)
tab(iris, Species, v2=Petal.Width)

# Petal.Width vs Species (will be different)
#   usage: tab(df, v1, v2=v2, nsize=FALSE)
tab(iris, Petal.Width, v2=Species)
```

## Weighted Tabs

To calculate weighted tabs, use the optional `weight_var` or `sdesign` parameters.  The first takes the name of a column with weights and the second takes a survey design object from the survey package.

Again, using the **iris** package even though it's not survey data:

```{r}
tab(iris, Species, weight_var = Petal.Length)
```

## Technical Notes

Full specification: 

```{r, eval=FALSE}
tab(df, v1, v2 = NULL, weight_var = NULL, sdesign = NULL, nsize = FALSE, ci = FALSE, to_factor=TRUE, collapses=list())
```

The function creates a temporary `vv1`, and if v2 is specified `vv2`, variable(s) in the data.  The function checks that this/these columns are not present before running, otherwise printing an error with no output.

The `stab` function (think s-tab) works the same way except that it takes quoted inputs.
```{r}
# s-tab version
stab(iris, "Species", v2="Petal.Width")

# equivalent tab version
tab(iris, Species, v2=Petal.Width)
```

The `tab` function wraps `stab`, which in turn wraps `utab` (for unweighted) and `wtab` (for weighted).

